---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Irene Chang, Susanna Jenkins, and Jordan Mullens"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()
setwd("/home/guest/Mullens_Chang_Jenkins_Final")

# Load your packages
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)

# Set your ggplot theme
library(ggthemes)
our_theme <- theme_base() +
  theme(
    line = element_line(
      color= 'black',
      linewidth = 2),
    legend.background = element_rect(
      color = "lightgrey",
      fill = "lightgrey"),
    legend.title = element_text (
      color = "blue"))
theme_set(our_theme)


# Load your datasets
RDD.raw <- read.csv("./Public_RDD_Spending_All_Countries.csv")


```


# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploration of Raw Data 

```{r dataset info}
#checking the original data set dimensions before wrangling. View dimensions, cloumn names, variable type, and head of each column.

dim(RDD.raw)
summary(RDD.raw)


```

# Data Wrangling
After the initial exploration, the dataset was wrangled to suit the analysis. The data frame was modified so that only the

```{r dataset wrangling}
#change column names, select 



```


\newpage

# Analysis

```{r}
#Format as a t-test
#EPAair$Ozone will be our continuous dependent variable
#EPAair$Year will be our categorical variable with two levels (2018 and 2019)
O3.twosample <- t.test(EPAair$Ozone ~ EPAair$Year)
O3.twosample

# Format as a GLM
O3.twosample2 <- lm(EPAair$Ozone ~ EPAair$Year)
summary(O3.twosample2)

```



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
